---
title: "NACA Stan fitting"
output: html_notebook
---

## Setup

```{r libraries, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(broom)
library(rstan)
library(gdata)
library(bayesplot)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

knitr::purl(input = "nacaAnalysis.Rmd",output = "nacaAnalysis.R")
```

## Simulation parameters

First thing to do is to set priors for parameters.
```{r priors}
prior_logkChemInt <- -3
priorSD_logkChemInt <- 0.5
prior_logkBleachInt <- -4
priorSD_logkBleachInt <- 0.5
prior_mChem <- -0.05
priorSD_mChem <- 0.5
prior_mBleach <- 0
priorSD_mBleach <- 0.5
prior_ext <- 0.02
priorSD_ext <- 0.01
prior_delay <- 2
priorSD_delay <- 0.01

priorSD_sigma <- 0.05

debug <- 0; #0 for no debug output, 1 for printing gen values, 2 for printing params

init_params <- list(list(logkChemInt=prior_logkChemInt,
                    logkBleachInt=prior_logkBleachInt,
                    mChem=prior_mChem,
                    mBleach=prior_mBleach,
                    ext=prior_ext,
                    delay=prior_delay,
                    sigma = 0.02))
init_params_4x <- rep(init_params,4)
```

Next, specify experimental conditions. These are in units of micromolar:
```{r conditions}
initCys <- 50
initDTP <- 100
```

Later we will fit data and so will have meaningful values for these data values. For now, we are just simulating data from the priors, so we will just set them to junk data
```{r nullData}
N <- 2
time <- c(5,6)
gdn <- c(0,0)
absorbance <- c(0,0)
numTrials <- 1
trialStarts <- c(1,3)
```

After estimating parameters, the model will use them to simulate data at [Gdn] = 0, 3, and 6. This specifies how many data points to simulate for each, as well as how long from mixing to the first data point recorded.
```{r nData}
nDat_gen <- 100
deadTime <- 5
```


Now we have to format all this data for stan.
```{r formatStanSim}
stan_data <-list(prior_logkChemInt=prior_logkChemInt, priorSD_logkChemInt=priorSD_logkChemInt,
                 prior_logkBleachInt=prior_logkBleachInt, priorSD_logkBleachInt=priorSD_logkBleachInt,
                 prior_mChem=prior_mChem, priorSD_mChem=priorSD_mChem, prior_mBleach=prior_mBleach,
                 priorSD_mBleach=priorSD_mBleach, prior_ext=prior_ext,priorSD_ext=priorSD_ext,
                 prior_delay=prior_delay, priorSD_delay=priorSD_delay, initCys=initCys, initDTP=initDTP, N=N,
                 time=time, gdn=gdn, absorbance=absorbance,
                 numTrials=numTrials,trialStarts=trialStarts,nDat_gen=nDat_gen,deadTime=deadTime,
                 priorSD_sigma=priorSD_sigma,debug=debug)
```

## Simulate data

Now we will simulate data from our priors using the stan file.
```{r stanSimulate}
stan_model<- "nacaModel.stan"
#stanc(file = stan_model, verbose = TRUE)
sim_data <- stan(file = stan_model,
             data = stan_data, init = init_params,
             chains = 1, iter = 1)
```

Plot the simulated data.
```{r plot-sim}
posterior <- rstan::extract(sim_data)

extractData <- function(fitObject,name,gdn){
  vals <- as.matrix(fitObject,pars = name)
  times <- as.matrix(fitObject, pars = "times_gen")
  vals <- as.vector(vals)
  times <- as.vector(times)
  pred_data <- data.frame("Time" = times, "Gdn" = gdn, "Abs" = vals)
  return(pred_data)
}
  
pred_data_0 <- extractData(sim_data,"pred_gdn_0",0)
pred_data_3 <- extractData(sim_data,"pred_gdn_3",3)
pred_data_6 <- extractData(sim_data,"pred_gdn_6",6)

pred_trial_starts = c(1,
                      nDat_gen+1,
                      nDat_gen*2+1,
                      nDat_gen*3+1)

numPredTrials = 3

pred_data_ALL <- rbind(pred_data_0,pred_data_3,pred_data_6)
ggplot(pred_data_ALL,aes(x = Time, y = Abs,color = factor(Gdn))) +
  geom_point() +
  labs(x = "Time (s)", y = "Absorbance", title = "Predicted Values") +
  scale_x_continuous(trans='log10', labels = scales::label_comma(), limits = c(1,1800))

```

## Fit simulated data (same priors)

Now, we try to go back and fit the simulated data with the model. First we need to format the data.

```{r format-sim}
pred_stan_data <- stan_data
# the same as before except for these changes
pred_stan_data$N <- 3*nDat_gen
pred_stan_data$time <- pull(pred_data_ALL,Time)
pred_stan_data$gdn <- pull(pred_data_ALL,Gdn)
pred_stan_data$absorbance <- pred_data_ALL %>%
  mutate(Abs = if_else(Abs<=0,0.0001,Abs)) %>%
  pull(Abs)
pred_stan_data$numTrials <- numPredTrials
pred_stan_data$trialStarts <- pred_trial_starts
```

First, we will fit using the same priors as the values used to generate the data.
```{r fit-same-priors}
sim_fit <- stan(file = stan_model,
             data = pred_stan_data, init = init_params_4x,
             chains = 4,iter = 2000)
```

Plot the posteriors, along with the center of the prior distributions.
```{r posterior-plots}
sim_fit_posterior <- rstan::extract(sim_fit)

par(mfrow = c(2,2))

plot(density(sim_fit_posterior$logkChemInt), main = "logkChemInt")
abline(v = prior_logkChemInt, col = 4, lty = 2)

plot(density(sim_fit_posterior$logkBleachInt), main = "logkBleachInt")
abline(v = prior_logkBleachInt, col = 4, lty = 2)

plot(density(sim_fit_posterior$mChem), main = "mChem")
abline(v = prior_mChem, col = 4, lty = 2)

plot(density(sim_fit_posterior$mBleach), main = "mBleach")
abline(v = prior_mBleach, col = 4, lty = 2)
```

Convergence plots: diagnostic to see fit.
```{r diagnostic-plots}
traceplot(sim_fit,pars=c("logkChemInt","logkBleachInt","mChem","mBleach"))
```

##Fit simulated data (different priors)

First, save old parameters as the target we hopefullly will get back to
```{r save-targets}
target_logkChemInt <- prior_logkChemInt
target_logkBleachInt <- prior_logkBleachInt
target_mChem <- prior_mChem
target_mBleach <- prior_mBleach
target_ext <- prior_ext
target_delay <- prior_delay
```

Now, change the priors
```{r change-priors}
prior_logkChemInt <- -2
priorSD_logkChemInt <- 2
prior_logkBleachInt <- -1
priorSD_logkBleachInt <- 2
prior_mChem <- 0.05
priorSD_mChem <- 2
prior_mBleach <- 0.1
priorSD_mBleach <- 2
prior_ext <- 0.01
priorSD_ext <- 0.02
prior_delay <- 5
priorSD_delay <- 1

priorSD_sigma <- 0.05

debug <- 2; #0 for no debug output, 1 for printing gen values, 2 for printing params

init_params <- list(list(logkChemInt=prior_logkChemInt,
                    logkBleachInt=prior_logkBleachInt,
                    mChem=prior_mChem,
                    mBleach=prior_mBleach,
                    ext=prior_ext,
                    delay=prior_delay,
                    sigma = 0.02))
init_params_4x <- rep(init_params,4)
```

Now format data as before
```{r format-sim-2}
pred_stan_data_2 <- pred_stan_data
# the same as the previous fit except for changing the priors

pred_stan_data_2$prior_logkChemInt <- prior_logkChemInt
pred_stan_data_2$priorSD_logkChemInt <- priorSD_logkChemInt
pred_stan_data_2$prior_logkBleachInt <- prior_logkBleachInt
pred_stan_data_2$priorSD_logkBleachInt <- priorSD_logkBleachInt
pred_stan_data_2$prior_mChem <- prior_mChem
pred_stan_data_2$priorSD_mChem <- priorSD_mChem
pred_stan_data_2$prior_mBleach <- prior_mBleach
pred_stan_data_2$priorSD_mBleach <- priorSD_mBleach
pred_stan_data_2$prior_ext <- prior_ext
pred_stan_data_2$priorSD_ext <- priorSD_ext
pred_stan_data_2$prior_delay <- prior_delay
pred_stan_data_2$priorSD_delay <- priorSD_delay
pred_stan_data_2$debug <- debug
```

Now fit the data
```{r fit-diff-priors}
sim_fit_diff <- stan(file = stan_model,
             data = pred_stan_data_2, init = init_params_4x,
             chains = 4,iter = 2000)
```

Posterior plots
```{r posterior-plots-2}
sim_fit_diff_posterior <- rstan::extract(sim_fit_diff)

par(mfrow = c(2,2))

plot(density(sim_fit_diff_posterior$logkChemInt), main = "logkChemInt",xlim = c(-3.1,-1.9))
abline(v = prior_logkChemInt, col = 4, lty = 2)
abline(v = target_logkChemInt, col = "red", lty = 2)

plot(density(sim_fit_diff_posterior$logkBleachInt), main = "logkBleachInt",xlim = c(-5,-0.9))
abline(v = prior_logkBleachInt, col = 4, lty = 2)
abline(v = target_logkBleachInt, col = "red", lty = 2)

plot(density(sim_fit_diff_posterior$mChem), main = "mChem",xlim = c(-0.08,0.06))
abline(v = prior_mChem, col = 4, lty = 2)
abline(v = target_mChem, col = "red", lty = 2)

plot(density(sim_fit_diff_posterior$mBleach), main = "mBleach",xlim = c(-5,0.12))
abline(v = prior_mBleach, col = 4, lty = 2)
abline(v = target_mBleach, col = "red", lty = 2)
```

Diagnostic plots
```{r diagnostic-plots-2}
traceplot(sim_fit_diff,pars=c("logkChemInt","logkBleachInt","mChem","mBleach"))
```


### Now fit actual data

First load in data. It was formatted usign Mathematica.

```{r input-data}
exp_data <- read_csv("NACAdata.csv")
trial_nums <- exp_data$Trial
exp_trial_starts <- match(unique(trial_nums),trial_nums)
exp_trial_starts <- append(exp_trial_starts,length(trial_nums)+1)
```

Format data
```{r format-exp}
exp_stan_data <- stan_data
# the same as before except for these changes
exp_stan_data$N <- nrow(exp_data)
exp_stan_data$time <- pull(exp_data,Time)
exp_stan_data$gdn <- pull(exp_data,Gdn)
exp_stan_data$absorbance <- exp_data %>%
  mutate(Abs = if_else(Abs<=0,0.0001,Abs)) %>%
  pull(Abs)
exp_stan_data$numTrials <- length(exp_trial_starts)-1
exp_stan_data$trialStarts <- exp_trial_starts
```

Choose priors
```{r choose-priors-exp}
prior_logkChemInt <- -2
priorSD_logkChemInt <- 2
prior_logkBleachInt <- -1
priorSD_logkBleachInt <- 2
prior_mChem <- -0.5
priorSD_mChem <- 2
prior_mBleach <- -2
priorSD_mBleach <- 2
prior_ext <- 0.01
priorSD_ext <- 0.02
prior_delay <- 5
priorSD_delay <- 1

priorSD_sigma <- 0.05

debug <- 2; #0 for no debug output, 1 for printing gen values, 2 for printing params

init_params <- list(list(logkChemInt=prior_logkChemInt,
                    logkBleachInt=prior_logkBleachInt,
                    mChem=prior_mChem,
                    mBleach=prior_mBleach,
                    ext=prior_ext,
                    delay=prior_delay,
                    sigma = 0.02))
init_params_4x <- rep(init_params,4)

exp_stan_data$prior_logkChemInt <- prior_logkChemInt
exp_stan_data$priorSD_logkChemInt <- priorSD_logkChemInt
exp_stan_data$prior_logkBleachInt <- prior_logkBleachInt
exp_stan_data$priorSD_logkBleachInt <- priorSD_logkBleachInt
exp_stan_data$prior_mChem <- prior_mChem
exp_stan_data$priorSD_mChem <- priorSD_mChem
exp_stan_data$prior_mBleach <- prior_mBleach
exp_stan_data$priorSD_mBleach <- priorSD_mBleach
exp_stan_data$prior_ext <- prior_ext
exp_stan_data$priorSD_ext <- priorSD_ext
exp_stan_data$prior_delay <- prior_delay
exp_stan_data$priorSD_delay <- priorSD_delay
```

Fit experimental data
```{r fit-exp}
exp_fit <- stan(file = stan_model,
             data = exp_stan_data, init = init_params_4x,
             chains = 4,iter = 2000)
```

Posterior plots
```{r posterior-plots-exp}
exp_fit_posterior <- rstan::extract(exp_fit)

par(mfrow = c(2,2))

plot(density(exp_fit_posterior$logkChemInt), main = "logkChemInt",xlim = c(-3.1,-1.9))
abline(v = exp_stan_data$prior_logkChemInt, col = 4, lty = 2)
plot(density(exp_fit_posterior$logkBleachInt), main = "logkBleachInt",xlim = c(-5,-0.9))
abline(v = exp_stan_data$prior_logkBleachInt, col = 4, lty = 2)
plot(density(exp_fit_posterior$mChem), main = "mChem",xlim = c(-0.2,0.06))
abline(v = exp_stan_data$prior_mChem, col = 4, lty = 2)
plot(density(exp_fit_posterior$mBleach), main = "mBleach",xlim = c(-5,0.12))
abline(v = exp_stan_data$prior_mBleach, col = 4, lty = 2)
```

Diagnostic plots
```{r diagnostic-plots-exp}
traceplot(exp_fit,pars=c("logkChemInt","logkBleachInt","mChem","mBleach"))
```



