---
title: "NACA Stan fitting"
output: html_notebook
---

```{r libraries, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(rstan)
library(gdata)
library(bayesplot)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


First thing to do is to set priors for parameters.
```{r priors}
prior_logkChemInt <- -3
priorSD_logkChemInt <- 0.5
prior_logkBleachInt <- -4
priorSD_logkBleachInt <- 0.5
prior_mChem <- -0.05
priorSD_mChem <- 0.5
prior_mBleach <- 0
priorSD_mBleach <- 0.5
prior_ext <- 0.02
priorSD_ext <- 0.01
prior_delay <- 2
priorSD_delay <- 0.01

priorSD_sigma <- 0.05

debug <- 0; #0 for no debug output, 1 for printing gen values, 2 for printing params

init_params <- list(list(logkChemInt=prior_logkChemInt,
                    logkBleachInt=prior_logkBleachInt,
                    mChem=prior_mChem,
                    mBleach=prior_mBleach,
                    ext=prior_ext,
                    delay=prior_delay,
                    sigma = 0.02))
init_params_4x <- rep(init_params,4)
```

Next, specify experimental conditions. These are in units of micromolar:
```{r conditions}
initCys <- 50
initDTP <- 100
```

Later we will fit data and so will have meaningful values for these data values. For now, we are just simulating data from the priors, so we will just set them to junk data
```{r nullData}
N <- 2
time <- c(5,6)
gdn <- c(0,0)
absorbance <- c(0,0)
numTrials <- 1
trialStarts <- c(1,3)
```

After estimating parameters, the model will use them to simulate data at [Gdn] = 0, 3, and 6. This specifies how many data points to simulate for each, as well as how long from mixing to the first data point recorded.
```{r nData}
nDat_gen <- 100
deadTime <- 5
```


Now we have to format all this data for stan.
```{r formatStanSim}
stan_data <-list(prior_logkChemInt=prior_logkChemInt, priorSD_logkChemInt=priorSD_logkChemInt,
                 prior_logkBleachInt=prior_logkBleachInt, priorSD_logkBleachInt=priorSD_logkBleachInt,
                 prior_mChem=prior_mChem, priorSD_mChem=priorSD_mChem, prior_mBleach=prior_mBleach,
                 priorSD_mBleach=priorSD_mBleach, prior_ext=prior_ext,priorSD_ext=priorSD_ext,
                 prior_delay=prior_delay, priorSD_delay=priorSD_delay, initCys=initCys, initDTP=initDTP, N=N,
                 time=time, gdn=gdn, absorbance=absorbance,
                 numTrials=numTrials,trialStarts=trialStarts,nDat_gen=nDat_gen,deadTime=deadTime,
                 priorSD_sigma=priorSD_sigma,debug=debug)
```


Now we will simulate data from our priors using the stan file.
```{r stanSimulate}
stan_model<- "nacaModel.stan"
#stanc(file = stan_model, verbose = TRUE)
sim_data <- stan(file = stan_model,
             data = stan_data, init = init_params,
             chains = 1, iter = 1)
```

Plot the simulated data.
```{r plot-sim}
posterior <- rstan::extract(sim_data)

extractData <- function(fitObject,name,gdn){
  vals <- as.matrix(fitObject,pars = name)
  times <- as.matrix(fitObject, pars = "times_gen")
  vals <- as.vector(vals)
  times <- as.vector(times)
  pred_data <- data.frame("Time" = times, "Gdn" = gdn, "Abs" = vals)
  return(pred_data)
}
  
pred_data_0 <- extractData(sim_data,"pred_gdn_0",0)
pred_data_3 <- extractData(sim_data,"pred_gdn_3",3)
pred_data_6 <- extractData(sim_data,"pred_gdn_6",6)

pred_trial_starts = c(1,
                      nDat_gen+1,
                      nDat_gen*2+1,
                      nDat_gen*3+1)

numPredTrials = 3

pred_data_ALL <- rbind(pred_data_0,pred_data_3,pred_data_6)
ggplot(pred_data_ALL,aes(x = Time, y = Abs,color = factor(Gdn))) +
  geom_point() +
  labs(x = "Time (s)", y = "Absorbance", title = "Predicted Values") +
  scale_x_continuous(trans='log10', labels = scales::label_comma(), limits = c(1,1800))

```

Now, we try to go back and fit the simulated data with the model. First we need to format the data.

```{r format-sim}
pred_stan_data <- stan_data
# the same as before except for these changes
pred_stan_data$N <- 3*nDat_gen
pred_stan_data$time <- pull(pred_data_ALL,Time)
pred_stan_data$gdn <- pull(pred_data_ALL,Gdn)
pred_stan_data$absorbance <- pred_data_ALL %>%
  mutate(Abs = if_else(Abs<=0,0.0001,Abs)) %>%
  pull(Abs)
pred_stan_data$numTrials <- numPredTrials
pred_stan_data$trialStarts <- pred_trial_starts

```

First, we will fit using the same priors as the values used to generate the data.
```{r fit-same-priors}
sim_fit <- stan(file = stan_model,
             data = pred_stan_data, init = init_params_4x,
             chains = 4,iter = 2000)
```

```{r extract-sim-fit}
sim_fit_posterior <- rstan::extract(sim_fit)
traceplot(sim_fit,pars=c("logkChemInt","logkBleachInt","mChem","mBleach"))
stan_dens(sim_fit,pars=c("logkChemInt","logkBleachInt","mChem","mBleach"))
```

